\documentclass[12pt]{article}
\include{commands}
\begin{document}\rztitle{Cognitive State Semantics and the Condition of the Sign}
\rzauthor{\[Name Witheld\]}{,}{[Affiliation Witheld]}
\begin{abstract}\noindent What are transcendental Conditions of Possibility for 
linguistics and semiotics to exist, the Ontology of the sign 
itself \mdash{} how that spoken, written, or inscribed mark differs from any other kind 
of being, so that it brings language into the world?  Language is a regime 
of multiple conversation partners, an ambient surrounding where they 
share actions, perceptions, and foci of attention, and a common posit of structured rules.  
If communication is successful, each sign's referent is 
isolated, collectively/phenomenologically, from a (canonically spatiotemporal) reality 
that extends beyond it.  I will review formal ideas 
(drawn, especially, from variants of Type Theory and Lambda Calculus) that 
sketch this process, but keeping in sight a cognitive nuance and intersubjectivity that is 
(thankfully) an essential, unavoidable part of human language. 
Concluding, I propose both cases for and limits on formal theories' 
applicability for Cognitive Linguistics/Humanities/Phenomenology, through the lens of the 
Philosophy of Science \mdash{} 
arguing how Phenomenology and Cognitive Linguistics, while distinct traditions, can work powerfully in consort.
\end{abstract}
\input{top-quotes}
\vspace{1em}
\p{A sign demands an individuation \mdash{} a criteriology, for anyone addressed and solicited by each sign, 
to recognize and isolate it as such.  Signs, and their referents, 
need an isolating, from the world around (and one another), 
or they are not signs.  In extreme cases a sign may 
stand alone, like the smoke fire telling a hiker's location; but by norm, 
embedded in discourse and performance, signs require (and carry with them, internally) 
understood inter-boundaries.  Both the recognition and the interpretation of signs therefore 
implicates cognition-logics of part/whole (mereology), and (dis/)continuity, 
each sign/referent disconnected in some ways, and continuous in others, 
with larger wholes inside which they are semi-autonomous parts.
}  
\p{Ad-hoc signs can blur the boundaries 
between conventionalized languages (verbal or not) and more impromptu social interactions 
\mdash{} the smoke fire is in part a conventional distress signal and in part a tool, an 
engineered natural phenomenon designed with intended effect, like causing rescuers to see it.  
The distress fire is also in a sense its own referent: its function as a sign is to call attention 
to itself, and so its location.  Or, choosing to write on one's own body 
\mdash{} like Hamas commander Mahmoud Ishtiwi, betrayed and killed by his own movement, carving 
into his leg the word \q{zulum} (\q{wronged}) \mdash{} is expression spreading beyond the  
conventions of words; the signs are made to signify the conditions of their execution.
These rather dramatic examples are more the provenance of semiotics than 
linguistics.  But people in ordinary verbal communication equally rely on a mixture of 
linguistic and other signs \mdash{} we point, make gestures, use \q{body language} and tone of voice.  
When conversation turns to some topic, like \q{that building over there}, such cues help 
speakers synchronize their attentions.  Tone and gestures clarify sentiment (honest, joking, sarcastic, anger) 
that may be ambiguous in spoken words alone, taken \q{out of context}.  
The weight of linguistic meaning is 
borne by semantics and pragmatics in fusion.  
Semantics has both formal and informal dimensions 
\mdash{} linking first to cognitive schema, or (as I will argue) prototypes of how 
schema are triggered; and second to pragmatics and contexts.
Conventionalized in semantic norms, schema, part abstract and part cognitive, 
help prime language users to manipulate formal structures 
in language, relative to the situational aether.  (Dis/)continuity in the plane of reference 
brings consciousness to a mereo-logic \cite{KitFine}, \cite{BarrySmith} that language-cognition 
can then reshape into syntax and semantics \cite{BittnerSmithDonnelly}.  
Semantic layers are abstract tools, but 
they offer a tableau of forms and combinations which users adapt, concretely, to each context.  
The deep potential of language, I believe, comes from the perpetual combination of the formal/abstract 
and the concrete/phenomenological.  
}
\p{For signs, the largest whole is a 
\q{plane} of articulation; for referents, it is an overall phenomenological surround; or, 
for more abstract signifieds, a space of concepts.  Like a footprint, whose very existence 
depends on both material continuity and visual break, for each sign there must be a blend of 
continuity and discontinuity, both around 
the sign and its referent.  Attending to a mereologically ordered world, 
we need innate theories warranting criteria for seeing things as both individuals 
and as causally/behaviorally constrained by and from a whole.  These criteria 
include structural consideration of the whole, and it is often in structural terms that 
the blend of autonomy and linkage for each part is realized.  Attunement to structured organization 
therefore warrants the perceptual and mental isolation of particular foci of attention 
\cite{WiegandMereology}, \cite{WiegandGestalts}. 
}
\p{As this plays out on planes of articulation alongside general situational awareness, the 
structures of discourse \mdash{} its division into distinct signs and their structural 
interrelationships \mdash{} and that of patterns we identify in our surroundings, that 
provide a context of discourse, play off one another.  Grammar does not iconify 
interrelationships among referents \mdash{} unlike  diagrams, maps, scientific simulations, 
or scale models \mdash{} but it ensures communicators may create structures among words that 
suggest, in each others' minds, concordant patterns in the environing world/situation.  
Even where there is direct sensory and perceptual evidence 
for objects' individuation and intercontinuity, visually and experientially present 
(which of course is only one kind of talk and reference), our preparedness to focus 
attention here or there depends on mental models of situations which are 
more abstract and schematic, and receptive to functional and interpersonal 
details.  The objects around us are not just blobs of matter, but usually 
have a constructed purpose, socially sanctioned 
meaning, nostalgic weight, and other significance that cannot be grasped by perception alone.
}
\p{A central theme in Cognitive Linguistics is that language meaning depends on situational 
understanding, and by extension on mental schema of spatial, temporal, and functional 
organization \mdash{} not only how environments are arranged, but how they are causally 
and physically determined \cite{Pinker}, \cite{Talmy}.  The difference between \i{pour water} and \i{spill water}, 
for example, is the person's deliberate intentions in relation to natural forces and 
tendencies (such as that of water to fall downward).  To \q{apply paint} to something, 
compared with to \q{cover} with paint, suggests different spatial configurations; to 
\i{fill a glass with water}, versus \i{pour water into} the glass, suggests both 
different spatial details and maybe rationale as well.  These are differences in 
emphasis, not necessarily in actuality: those pairs of alternatives could describe 
identical state of affairs. \hspace{-2pt} But they direct conversants' attention in different ways, they 
choose one or another part of a scene as a reference frame, and suggest 
different \q{takes}.  These are driven by \i{semantic} variations 
\mdash{} the choice of verbs like \i{pour} versus \i{fill}, \i{pour} versus \i{spill}, 
or \i{apply} versus \i{cover}.  But semantic and syntactic rules work in federation, 
relative to context: for example, different verbs take different prepositions in different 
situations.  Pour \i{into} vs. fill \i{with}.  To join \q{pour} 
with \i{with} places emphasis elsewhere \mdash{} onto the device which enables the 
pourer to do the pouring.  So the grammatic and semantic norms of a language 
jointly offer a terrain of options from which speakers assemble combinations 
invoking those aspects of situations that they wish to emphasize. 
}
\p{In short, grammar is language's substitute for \i{visual} or \i{physical} resemblance-to-structure.  
Take Ronald Langacker's \q{landmark}/\q{trajector} model as an example: one (very general) 
manner of spatial gestalt, subject to either intuitive, reflective analysis or to formalization 
(\cite{BergenChang}, \cite{VisettiCadiot}).
\q{That boat crossing the lake}: \i{boat} (trajector) 
perceived against \i{lake} (landmark), which provides context; together they 
produce a mental model; a figured spatial relationship.  This is 
communicated, not by visual or kinaesthetic effect,\footnote{At least in prose \mdash{} poetry, which can bring back a semiotics of raw visual layout and auditory effect, is an 
exception that proves the rule.
} but by the more abstract effects of intentions signaled, 
via both exact words (\q{crossing} paints a different picture than would \i{across}, \i{on}, \i{by}; 
still more so, \i{at the bottom of}), and morphosyntactic tropes 
(like the form \XRelY{}; where \q{\Rel{}} here means one from many spatial relations, taking 
\i{trajector} to the left and \i{landmark} to the right).  \i{Landmark} and \i{trajector} are anchors around which 
both syntactic and semantic selections are organized.
}
\p{Language needs both abstract laws and cognitively-mediated construals of ambient situations.  
The abstract laws are shaped by the situations, not directly \mdash{} it is that extra 
indirection which cleaves language from other sign systems \mdash{} 
but derivatively: language rules are optimized for conversants to 
mold linguistic possibilities into selection-spaces, which then become 
raw materials for representations of 
situational context.  Each choice of word and form adds a piece to 
a representational complex, and the sum of those pieces \mdash{} be this a sentence, 
a conversation turn, or an entire discourse \mdash{} is a language act that hews 
to the structure of a situation, as the speaker wants to emphasize it.  
Here I take this perspective as pregiven: not as a perfected or homogenous theory, 
but as a working hypothesis on the origins of linguistic structuration as 
such.  The scope of this paper is then to analyze its ramifications for our understanding 
of grammar and formal semantics.
}
\input{section1.ngml}
\input{section2.ngml}
\input{section3.ngml}
\input{biblio}
\end{document}
