
%\vspace{.4em}

%\onecolumngrid
\onecolumn
\begin{thebibliography}{99}
	
\vspace{-.7em}

\bibitem{AbadiCardelli}
Mart\'{\OldI}n Abadi and Luca Cardelli,
\bibtitle{A Theory of Primitive Objects: 
Untyped and First-Order Systems}
\tinyurl{http://lucacardelli.name/Papers/PrimObj1stOrder.pdf}

%
\bibitem{RaubalAdams}
Benjamin Adams and Martin Raubal, 
\bibtitle{A Metric Conceptual Space Algebra}.
\tinyurl{https://pdfs.semanticscholar.org/521a/cbab9658df27acd9f40bba2b9445f75d681c.pdf}

\bibitem{RaubalAdamsCSML}
Benjamin Adams and Martin Raubal, 
\bibtitle{Conceptual Space Markup Language (CSML): Towards the Cognitive Semantic Web}.
\tinyurl{http://idwebhost-202-147.ethz.ch/Publications/RefConferences/ICSC_2009_AdamsRaubal_Camera-FINAL.pdf}

\bibitem{StAmant}
Robert St. Amant, \i{et. al.}, 
\bibtitle{An Image Schema Language}
\tinyurl{https://apps.dtic.mil/dtic/tr/fulltext/u2/a458943.pdf}


\bibitem{AkbikBross}
Alan Akbik and J\"urgen Bro\ss,
\bibtitle{Wanderlust: Extracting Semantic Relations from Natural
Language Text Using Dependency Grammar Patterns}
\tinyurl{http://ceur-ws.org/Vol-491/semse2009_7.pdf}

	
%\bibitem{Allam}
%Diana Allam, R\'emi Douence, Herv\'e Grall, Jean-Claude Royer, and Mario S\"udholt, 
%\bibtitle{Well-Typed Services Cannot Go Wrong}. [Research Report] RR-7899, 2012. hal-00700570v1.
%\tinyurl{https://hal.inria.fr/hal-00700570v1/document}

\bibitem{AnounEtAl}
Houda Anoun, Pierre Cast\'eran, and Richard Moot, 
\bibtitle{Proof Automation for Type-Logical Grammars}
\tinyurl{http://www.labri.fr/perso/moot/esslli_reader.pdf}


%??
\bibitem{AsherPustejovsky}
Nicholas Asher and James Pustejovsky, 
\bibtitle{A Type Composition Logic for Generative Lexicon}
\tinyurl{https://www.cs.brandeis.edu/~jamesp/classes/cs216-2009/readings2009/TCLforGL.pdf}



%
\bibitem{BagetEtAl}
Jean-Franc\c{o}is Baget \i{et. al}, 
	\bibtitle{Translations between RDF(S) and Conceptual Graphs}.
	\tinyurl{http://www.lirmm.fr/~croitoru/rdfs.pdf}	
	
%?
\bibitem{BarkerShan}
Chris Barker and Chung-Chieh Shan, 
\bibtitle{Continuations and Natural Language}.
\tinyurl{http://semanticsarchive.net/barker/barker-shan-continuations-book.pdf}

%
%\bibitem{Shan}
%Chris Barker and Chung-Chieh Shan, 
%\bibtitle{Continuations and Natural Language}.
%\tinyurl{http://semanticsarchive.net/barker/barker-shan-continuations-book.pdf}

%
\bibitem{Belohlavek}
Radim B\v{e}lohl\'avek and Vladim{\i\OldI}r Sklen\'a\v{r}, 
\bibtitle{Formal Concept Analysis Constrained by Attribute-Dependency Formulas}
B. Ganter and R. Godin, eds.,: ICFCA 2005, LNCS 3403, pp. 176-191, Berlin, Springer-Verlag, 2005.
\tinyurl{http://belohlavek.inf.upol.cz/publications/BeSk_Fcacadf.pdf}


%
%\bibitem{Benovsky}
%Juri Benovsky, 
%\bibtitle{A modal bundle theory}.  Metaphysica, 7.2, 2006.

\bibitem{BergenChang}
Benjamin K. Bergen and Nancy Chang, 
\bibtitle{Embodied Construction Grammar 
in Simulation-Based Language Understanding}.
\tinyurl{http://www1.icsi.berkeley.edu/~nchang/pubs/ecg.pdf}

\bibitem{BernardyEtAl}
Jean-Philippe Bernardy, et. al.,
\bibtitle{Parametricity and Dependent Types}.
\tinyurl{http://www.staff.city.ac.uk/~ross/papers/pts.pdf}

\bibitem{BittnerSmithDonnelly}
Thomas Bittner, Barry Smith, and Maureen Donnelly,
\bibtitle{The logic of systems of granular partitions.}
\tinyurl{http://ontology.buffalo.edu/smith/articles/BittnerSmithDonnelly.pdf}


\bibitem{BoasSag}
Hans C. Boas and Ivan A. Sag, eds. 
\bibtitle{Sign-Based Construction Grammar}.
\tinyurl{http://lingo.stanford.edu/sag/papers/theo-syno.pdf}

\bibitem{MathieuBouchard}
Mathieu Bouchard, 
\bibtitle{A Type Theory for the Documentation of PureData}
\tinyurl{http://artengine.ca/~catalogue-pd/43-Bouchard.pdf}

%
\bibitem{LineBrandt}
Line Brandt, 
\bibtitle{The Communicative Mind.}
Newcastle-upon-Tyne: Cambridge Scholars Publishing, 2013

%
\bibitem{PerAageBrandt}
Per Aage Brandt, 
\bibtitle{Spaces, Domains, and Meaning: Essays in Cognitive Semiotics}.
Newcastle-upon-Tyne: Cambridge Scholars Publishing, 2013
\tinyurl{http://semiotics.au.dk/fileadmin/Semiotics/pdf/per-aage-brandt/Per_Aage_Brandt__Spaces__Domains__and_Meaning._Essays_in_Co.pdf}


\bibitem{CasatiVarzi}
Roberto Casati and Achille C. Varzi, 
\bibtitle{Parts and Places: The Structures of Spatial Representation}
MIT Press, 1999


%\bibitem{MelChen}
%\bibtitle{Speech Lost from Speech: The Cognitive Linguistics of Alienation, Objectification, and Reclaiming}.
%Doctoral Dissertation, Berkeley.
%\tinyurl{http://escholarship.org/uc/item/7rv3j4nf}

\bibitem{ChatzikyriakidisLuo}
Stergios Chatzikyriakidis and Zhaohui Luo, 
\bibtitle{
	Individuation Criteria, Dot-types and Copredication:
	A View from Modern Type Theories.}
\intitle{Association for Computational Linguistics, Proceedings of the 14th 
Meeting on the Mathematics of Language (MoL 14)}, pp. 39–50, 2015. 
\tinyurl{http://www.aclweb.org/anthology/W15-2304}

\bibitem{NathanielChristenCyberphysical}
Nathaniel Christen, 
\bibtitle{Hypergraph Type Theory for 
Specifications-Conformant Code and 
Generalized Lambda Calculus}.
\intitle{Advances in Ubiquitous Computing: Cyber-Physical Systems, Smart Cities, and Ecological Monitoring}, Amy Neustein, ed., Elsevier, 2019. 


\bibitem{MaryDalrymple}
Mary Dalrymple, \bibtitle{Lexical Functional Grammar}.
\tinyurl{http://web.stanford.edu/group/nasslli/courses/as-cr-da/apbook-nasslli.pdf}


\bibitem{Debusmann}
Ralph Debusmann, 
\bibtitle{An Introduction to Dependency Grammar}.
Hauptseminar \intitle{Dependenzgrammatik}, SoSe 99.
\tinyurl{https://www.ps.uni-saarland.de/~rade/papers/dg.pdf}

\bibitem{DebusmannThesis}
\underlines
\bibtitle{Extensible Dependency Grammar:
A Modular Grammar Formalism Based On
Multigraph Description.}
Universit\"at des Saarlandes, dissertation 2006.

\bibitem{DebusmannDuchierRossberg}
Ralph Debusmann, Denys Duchier and Andreas
Rossberg, 
\bibtitle{Modular Grammar Design with
	Typed Parametric Principles}.
James Rogers, ed., 
\intitle{Formal Grammar/Mathematics of Language 2005}, CSLI Publications, 2009.
\tinyurl{http://web.stanford.edu/group/cslipublications/cslipublications/FG/2005/debusmann.pdf}.

%
\bibitem{DelpeuchPreller}
Antonin Delpeuch and Anne Preller, \bibtitle{From Natural Language to RDF Graphs with Pregroups}.
\intitle{Proceedings of the EACL 2014 Workshop on Type Theory and Natural Language Semantics (TTNLS)}, pp. 55–62,
Gothenburg, Sweden, April 26-30 2014. 
\tinyurl{https://www.aclweb.org/anthology/W/W14/W14-1407.pdf}


\bibitem{Descles2010}
Jean-Pierre Descl\'es, \bibtitle{Reasoning in Natural Language in 
	Using Combinatory Logic and Topology: An Example with Aspect
	and Temporal Relations.}  
\tinyurl{https://www.aaai.org/ocs/index.php/FLAIRS/2010/paper/viewFile/1350/1736}


\bibitem{Escardo}
Mart{\'\OldI}n Escard\'o, \bibtitle{Synthetic topology of data types and classical spaces}.  
\intitle{Electronic Notes in Theoretical Computer Science} 87, 2004. 
\tinyurl{http://www.cs.bham.ac.uk/~mhe/papers/barbados.pdf}


\bibitem{EyreLawry}
Henrietta Eyre and Jonathan Lawry,
\bibtitle{Language Games with Vague Categories and Negations}
\tinyurl{http://adb.sagepub.com/content/early/2014/09/04/1059712314547318}


\bibitem{Fauconnier}
Gilled Fauconnier, \bibtitle{Mental Spaces: Aspects of Meaning Construction in Natural Language}.
Cambridge, 1994

\bibitem{KitFine}
Kit Fine, \bibtitle{Part-whole}, in Barry Smith and 
David Woodruff Smith, \intitle{The Cambridge Companion to Husserl}, Cambridge University Press, New York, 1995, 
pp. 463.
	
%\bibitem{Gangle}
%Rocco Gangle, \bibtitle{Diagrammatic Immanence: Category Theory and Philosophy}.  
%Edinburgh University Press, 2015	

\bibitem{GaumeVenantVictorri}
Bruno Gaume, Fabienne Venant, and Bernard Victorri,  
\bibtitle{Hierarchy in Lexical Organisation of Natural Languages}.
\intitle{Hiearchy in Natural and social Sciences}, 
Methodos series, vol 3, Springer, 2006, pp. 1039-1058, 2013.  
\tinyurl{https://pdfs.semanticscholar.org/ac47/62895ba5019bd556a8c9b4cde0a36b20ea0c.pdf}

\bibitem{Zenker}
Peter \Gardenfors{} and Frank Zenker,  
\bibtitle{Theory Change as Dimensional Change: Conceptual Spaces 
	Applied to the Dynamics of Empirical Theories}.
\intitle{Synthese 190(6)}, pp. 1039-1058, 2013.  
\tinyurl{http://lup.lub.lu.se/record/1775234}

\bibitem{GiorgoloAsudeh}
Gianluca Giorgolo and Ash Asudeh, \bibtitle{Monads as a Solution for Generalized Opacity}.
Proceedings of the EACL 2014 Workshop on Type Theory and Natural Language Semantics (TTNLS), pages 19–27,
Gothenburg, Sweden, April 26-30 2014. 
\tinyurl{http://www.aclweb.org/anthology/W14-1403}


\bibitem{GoguenWhatIs}
Joseph Goguen, \bibtitle{What is a Concept?}
\tinyurl{https://cseweb.ucsd.edu/~goguen/pps/iccs05.pdf}


\bibitem{GoguenSemioticMorphism}
\underlines \bibtitle{Semiotic Morphisms, Representations, and Blending for 
	Interface Design}
\tinyurl{https://cseweb.ucsd.edu/~goguen/pps/uid.pdf}

\bibitem{AlbertGoldfain}
Albert Goldfain, \bibtitle{Vital Sign Ontology}.
\tinyurl{https://philarchive.org/archive/GOLVSOv1}

\bibitem{HongEisner}
Juneki Hong and Jason Eisner,
\bibtitle{Deriving Multi-Headed Planar Dependency Parses
from Link Grammar Parses}
\tinyurl{https://www.cs.jhu.edu/~jason/papers/hong+eisner.tlt14.paper.pdf}


\bibitem{KennethHolmqvist}
Kenneth Holmqvist,
\bibtitle{Conceptual Engineering: Implementing cognitive semantics}, 
in Jens Allwood and Peter \Gardenfors, eds., 
\intitle{Cognitive Semantics}, pp 153 - 171, Amsterdam,
Philadelphia: John Benjamins, 1999.

\bibitem{HolmqvistDiss}
Kenneth Holmqvist, 
\bibtitle{Implementing Cognitive Semantics: Image schemata, 
valence accommodation, and valence suggestion for  
AI and computational linguistics}.
PhD thesis, Dept. of Cognitive Science, Lund University, Lund, Sweden, 1993.

\bibitem{HOTT}
The Univalent Foundations Program,
\bibtitle{Homotopy Type Theory: Univalent Foundations of Mathematics}
Princeton 2013
\tinyurl{https://hott.github.io/book/nightly/hott-online-1075-g3c53219.pdf}


%\bibitem{Husserl}
%Edmund Husserl, \bibtitle{Experience and Judgment.}
%James S. Churchill and Karl Ameriks, tr.    
%Northwestern, 1973.


\bibitem{Johnson}
Mark Johnson, \bibtitle{The Body in the Mind: 
The Bodily Basis of Meaning, Imagination, and Reason}.  %Chicago, 1990
University of Chicago Press, 1990


%\bibitem{Jubien}
%Michael Jubien, 
%\bibtitle{Ontology, Modality, and the Fallacy of Reference}.
%Cambridge University Press, 2009.

%\bibitem{HyderRheinberger}
%David Hyder and Hans-J\"org Rheinberger, eds.,
%\bibtitle{Science and the Life-World: Essays on Husserl's \underline{Crisis of European Sciences}.}
%Stanford University Press, 2009.

\bibitem{SeanDorranceKelley}
Sean Dorrance Kelley, \bibtitle{Seeing Things in Merleau-Ponty}.
\tinyurl{http://www.nyu.edu/gsas/dept/philo/courses/representation/papers/Kelly.pdf}

\bibitem{KepserMonnich}
Stephan Kepser and Uwe M\"onnich, 
\bibtitle{(Un-)Decidability Results for Head-Driven Phrase Structure
	Grammar}.
\itemtitle{Algebraic Methods in Language Processing AMiLP-3},
Anton Nijholt and Giuseppe Scollo, eds., 2003.
\tinyurl{http://www.tcl-sfs.uni-tuebingen.de/kepser/papers/amilp3-hpsg.pdf}

\bibitem{OzkanKilic}
Ozkan Kilic,
\bibtitle{Intelligent Reasoning on Natural Language Data:
A non-axiomatic reasoning system approach}
\tinyurl{https://pdfs.semanticscholar.org/2324/87388d5a11ec729e352501bd0012e89cf7de.pdf}

\bibitem{KiselyovControl}
Oleg Kiselyov, 
\bibtitle{How to Remove a Dynamic Prompt: Static and dynamic delimited continuation operators are equally expressible}
\tinyurl{https://www.researchgate.net/publication/238123187_How_to_remove_a_dynamic_prompt_static_and_dynamic_delimited_continuation_operators_are_equally_expressible}

\bibitem{Kiselyov}
Oleg Kiselyov, 
\bibtitle{Applicative Abstract Categorial Grammars}.
\tinyurl{http://okmij.org/ftp/gengo/applicative-symantics/AACG.pdf}

\bibitem{KiselyovShan}
Oleg Kiselyov and Chung-chieh Shan,
\bibtitle{Lambda: the ultimate syntax-semantics interface}
\tinyurl{https://pdfs.semanticscholar.org/fb41/793ae7098c40fcdb6706b905c48a270b0b48.pdf}

%\bibitem{Kohn}
%Arne K\"ohn and Wolfgang Menzel,
%\bibtitle{Incremental Predictive Parsing with TurboParser}.
%\tinyurl{http://edoc.sub.uni-hamburg.de/informatik/volltexte/2014/203/pdf/koehn_incremental_predictive_parsing_with_turboparser_korrigiert.pdf}

\bibitem{KongRushSmith}
Lingpeng Kong, Alexander M. Rush, and Noah A. Smith, 
\bibtitle{Transforming Dependencies into Phrase Structures}.
\tinyurl{http://www.aclweb.org/anthology/H01-1014}

%???
%\bibitem{AlexanderKravchenko}
%Alexander V. Kravchenko,
%\bibtitle{Grammar as semiosis and cognitive dynamics}.
%\tinyurl{http://www.univie.ac.at/constructivism/archive/fulltexts/480.pdf}

\bibitem{LakoffJohnson}
George Lakoff and Mark Johnson, 
\bibtitle{Philosophy in the Flesh: the Embodied Mind and its Challenge to Western Thought}.
New York: Basic Books, 1999.

\bibitem{Langacker87}
Ronald Langacker, 
\bibtitle{Nouns and Verbs}
\intitle{Language}, Vol. 63, No. 1 (March 1987) pp. 53-94.

\bibitem{LangackerFoundations}
Ronald Langacker, 
\bibtitle{Foundations of Cognitive Grammar, vol. 1}.  
Stanford University Press, 1991

\bibitem{AntonioLieto}
\bibtitle{Dual PECCS: A Cognitive System for Conceptual 
Representation and Categorization}
\tinyurl{https://www.researchgate.net/publication/304208795_Dual_PECCS_A_Cognitive_System_for_Conceptual_Representation_and_Categorization}


\bibitem{ZhaohuiLuo}
Zhaohui Luo,
\bibtitle{Type-Theoretical Semantics with Coercive Subtyping}.
\tinyurl{https://www.cs.rhul.ac.uk/home/zhaohui/ESSLLI11notes.pdf}

\bibitem{LuoSoloviev}
Zhaohui Luo and Sergei Soloviev,
\bibtitle{Dependent Coercions}.
\tinyurl{https://www.sciencedirect.com/science/article/pii/S1571066105803147}

\bibitem{MaillardClarkGrefenstette}
Jean Maillard, Stephen Clark, and Edward Grefenstette,
\bibtitle{A Type-Driven Tensor-Based Semantics for CCG}.
\tinyurl{http://www.cl.cam.ac.uk/~sc609/pubs/eacl14types.pdf}

\bibitem{MartinPollard}
Scott Martin and Carl Pollard,
\bibtitle{A Dynamic Categorial Grammar}.
\tinyurl{http://www.ling.ohio-state.edu/~murat/dycg.pdf}

\bibitem{MeryMootRetore}
Bruno Mery, Richard Moot, and Christian Retor\'e, \bibtitle{
	Plurals: Individuals and sets
	in a richly typed semantics}.  
\tinyurl{http://arxiv.org/pdf/1401.0660.pdf} \archiveDate{3 Jan 2014}

\bibitem{MeryRetore}
Bruno Mery and Christian Retor\'e, \bibtitle{
	Semantic Types,
Lexical Sorts and Slassifiers}. 
\intitle{10th International Workshop on
Natural Language Processing and Cognitive Science (NLPCS '10), 15
October 2013 - 17 October 2013 (Marseilles, France)} 
\tinyurl{https://core.ac.uk/download/pdf/78384461.pdf}

\bibitem{DavidMorris}
David Morris,
\bibtitle{The Sense of Space}. State University of New York, 2004

\bibitem{MullerPDF}
Stefan M\"uller, 
\bibtitle{Unifying everything: Some remarks on simpler syntax, construction grammar,
	minimalism, and HPSG}
\tinyurl{http://www.linguisticsociety.org/sites/default/files/Lg_89_4_Muller.pdf}

\bibitem{MullerBook}
Stefan M\"uller, 
\bibtitle{Grammatical theory: From transformational grammar to constraint-based approaches}.
Berlin: Language Science Press, 2016


\bibitem{Nivre}
Joakim Nivre, \bibtitle{Dependency Grammar and Dependency Parsing.}
\tinyurl{http://stp.lingfil.uu.se/~nivre/docs/05133.pdf}

\bibitem{EstherPascual}
Esther Pascual, 
\bibtitle{Fictive Interaction: The conversation frame in thought, language, and discourse}.
Philadelphia, John Benjamins, 2014.

\bibitem{SantanuPaul}
Santanu Paul and Atul Prakash,
\bibtitle{Supporting Queries on Source Code: A Formal Framework}
\tinyurl{https://www.cse.umich.edu/techreports/cse/94/CSE-TR-209-94.pdf}


\bibitem{EfePeker}
Efe Peker,
\bibtitle{Following 9/11: George W. Bush's Discursive Re-Articulation of American Social Identity}
Master's Dissertation, Link\"oping University, Sweden
\tinyurl{http://www.diva-portal.org/smash/get/diva2:21407/FULLTEXT01.pdf}


\bibitem{NaturalizingPhenomenology}
Jean Petitot, et. al., eds., \bibtitle{Naturalizing Phenomenology: 
	Issues in Contemporary Phenomenology and Cognitive Science}.
Stanford University Press, 1999.

\bibitem{PetitotSyntaxe}
Jean Petitot, \bibtitle{Syntax Topologique et Grammaire Cognitive}.
\intitle{Langages 25.103}, pp. 97-128, 1991. 

\bibitem{JeanPetitot}
Jean Petitot, \bibtitle{The morphodynamical turn of cognitive linguistics}.
\tinyurl{https://journals.openedition.org/signata/549}


\bibitem{Pinker}
Steven Pinker, \bibtitle{The Stuff of Thought:  Language As a Window Into Human Nature}.
Penguin, 2007.  

%\bibitem{GriseldaPollock}
%Griselda Pollock, \bibtitle{Modernity and the Spaces of Femininity}.
%\tinyurl{https://msu.edu/course/ha/446/griseldapollock.pdf}

\bibitem{Pollard}
Carl Pollard, \bibtitle{Covert Movement in Logical Grammar}.
\tinyurl{http://www.ling.ohio-state.edu/~pollard/cvg/covert.pdf}


\bibitem{MartinRaubal}
Martin Raubal,
\bibtitle{Formalizing Conceptual Spaces}.
\tinyurl{http://www.raubal.ethz.ch/Courses/288MR_Spring08_Papers/Raubal_FormalizingConceptualSpaces_FOIS04.pdf}


%\bibitem{Rapaport}
%William J. Rapaport,
%\bibtitle{How to Pass a Turing Test: 
%Syntactic Semantics, Natural-Language Understanding,
%and First-Person Cognition}.
%Journal of Logic, Language, and Information 9: pp. 467-490, 2000. 
%\tinyurl{http://www.cse.buffalo.edu/~rapaport/Papers/TURING.pdf}

\bibitem{GiuseppeRiva}
Giuseppe Riva and Fabrizia Mantovani, 
\bibtitle{Being There: Understanding the Feeling of Presence in a Synthetic Environment
	and Its Potential for Clinical Change}.
\tinyurl{http://cdn.intechopen.com/pdfs-wm/39042.pdf}

\bibitem{Rossi}
\Aurelie{} Rossi, \bibtitle{Applicative and Combinatory Categorial Grammar:
	Analysis of the French Interrogative Sentences}.
\intitle{FLAIRS Conference, Association for the Advancement of Artificial
Intelligence}, 2008.
\tinyurl{https://www.aaai.org/Papers/FLAIRS/2008/FLAIRS08-118.pdf}

\bibitem{StellConvexivity}
Anthony J. Roy and John Stell, \bibtitle{Convexivity in Discrete Space}.  
\tinyurl{http://www.comp.leeds.ac.uk/jgs/RoyStellCOSIT03.pdf}

\bibitem{Saghafi}
Salman Saghafi, et. al., \bibtitle{Features and Object Capabilities: Reconciling Two Visions of Modularity}.
\tinyurl{https://cs.brown.edu/~sk/Publications/Papers/Published/sfk-feat-ocap-reconcil/paper.pdf}

\bibitem{ShanMonads}
Chung-Chieh Shan,
\bibtitle{Monads for natural language semantics}.
\tinyurl{http://arxiv.org/pdf/cs/0205026.pdf} \archiveDate{17 May 2002}

\bibitem{ShanThesis}
\underlines
\bibtitle{Linguistic side effects.}
Harvard University, dissertation 2005. 

\bibitem{Schneider}
Gerold Schneider, 
\bibtitle{A Linguistic Comparison of Constituency, Dependency and Link Grammar}.
Zurich University, diploma, 2008.
\tinyurl{https://files.ifi.uzh.ch/cl/gschneid/papers/FINALSgeroldschneider-latl.pdf}

\bibitem{MattSelway}
Matt Selway, \i{et. al},
\bibtitle{Configuring Domain Knowledge for Natural Language Understanding}
\tinyurl{http://ceur-ws.org/Vol-1128/paper9.pdf}


\bibitem{ShanControl}
Chung-Chieh Shan, 
\bibtitle{Shift to control}.
Proceedings of the \i{Workshop on Scheme and
Functional Programming`}, pp. 99–107, 2004.
\tinyurl{http://homes.sice.indiana.edu/ccshan/recur/recur.pdf}

\bibitem{SleatorTamperley}
Daniel D. Sleator and Davy Tamperley, 
\bibtitle{Parsing English with a Link Grammar}.
\tinyurl{https://www.link.cs.cmu.edu/link/ftp-site/link-grammar/LG-IWPT93.pdf}

\bibitem{BarrySmith}
Barry Smith, \bibtitle{Mereotopology: A Theory of Parts and
Boundaries}.
\tinyurl{http://nemo.nic.uoregon.edu/wiki/images/3/38/Smith1996_Mereotopology.pdf}.

\bibitem{BarrySmithBlood}
Barry Smith, \bibtitle{The Ontology of Blood Pressure: A Case Study in Creating Ontological Partitions in Biomedicine}.
\tinyurl{https://www.researchgate.net/publication/228961604_The_Ontology_of_Blood_Pressure_A_Case_Study_in_Creating_Ontological_Partitions_in_Biomedicine}

\bibitem{DavidWoodruffSmith}
David Woodruff Smith,
\bibtitle{Mind World}.  Cambridge University Press, 2004.

\bibitem{DavidWoodruffSmithChapter}
\underlines \bibtitle{Mind and body}, \bibtitle{The Cambridge Companion to Husserl}, New York 1995.

\bibitem{SmithMcIntyre}
David Woodruff Smith and Ronald McIntyre, \bibtitle{Husserl and Intentionality: 
A Study of Mind, Meaning, and Language}.
Dordrecht and Boston: D. Reidel, 1982.

\bibitem{SmithGibbons}
Michael Anthony Smith and Jeremy Gibbons,
\bibtitle{Unifying Theories of Objects}
\tinyurl{http://www.cs.ox.ac.uk/jeremy.gibbons/publications/uto.pdf}

%\bibitem{SowaMulti}
%John F. Sowa, 
%\bibtitle{Laws, Facts, and Contexts:
%Foundations for Multimodal Reasoning}.
%\tinyurl{http://www.jfsowa.com/pubs/laws.htm}

\bibitem{Sowa}
\underlines 
\bibtitle{From Existential Graphs to Conceptual Graphs}.
\tinyurl{http://www.jfsowa.com/pubs/eg2cg.pdf}

%\bibitem{Spivak}
%David Spivak,
%\bibtitle{Database Queries and Constraints via Lifting Problems}.
%\tinyurl{http://math.mit.edu/~dspivak/informatics/LiftingProblems.pdf}

\bibitem{SpivakKent}
David Spivak and Robert Kent, 
\bibtitle{Ologs: A Categorical Framework for Knowledge
Representation}
\tinyurl{https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0024274&type=printable}

\bibitem{Strle}
Gregor Strle, \bibtitle{Semantics Within: The Representation of Meaning Through Conceptual Spaces}.  
Univ. of Novi Gorici, dissertation, 2012.


\bibitem{Talmy}
Leonard Talmy, \bibtitle{Force Dynamics in Language and Cognition}.
From \bibtitle{Toward a Cognitive Semantics}, Vol. 1, MIT Press 2000.
\tinyurl{http://linguistics.buffalo.edu/people/faculty/talmy/talmyweb/Volume1/chap7.pdf}


\bibitem{TanakaEtAl}
Ribeka Tanaka, et. al., \bibtitle{Factivity and Presupposition
in Dependent Type Semantics}
\tinyurl{http://www.lirmm.fr/tytles/Articles/Tanaka.pdf}

\bibitem{ElkeTeich}
Elke Teich,
\bibtitle{Systemic Functional Grammar \& Natural Language Generation}
New York: Cassell, 1999.

%??
%\bibitem{Toninho}
%Bernardo Toninho, et. al., 
%\bibtitle{Functions as Session-Typed Processes}.
%\tinyurl{http://www.cs.cmu.edu/~btoninho/fossacs12.pdf}

\bibitem{OrlinVakarelov}
Orlin Vakarelov,
\bibtitle{Pre-cognitive Semantic Information}.
\tinyurl{https://link.springer.com/article/10.1007/s12130-010-9109-5}

\bibitem{VakarelovAgent}
Orlin Vakarelov, 
\bibtitle{The cognitive agent: Overcoming
informational limits}
\tinyurl{https://philarchive.org/archive/VAKTCAv1}

\bibitem{LaureVieu}
\bibtitle{Representing Content: Semantics, Ontology, and their Interplay}
Dissertation, Universit\'e Paul Sabatier, Toulouse
\tinyurl{https://www.irit.fr/publis/LILAC/LV-HDR09.pdf}

\bibitem{AnneVilnat}
Anne Vilnat,
\bibtitle{Dialogue et analyse de phrases}
Habilitation M\'emoire, Universit\'e Paris-Sud  
\tinyurl{https://perso.limsi.fr/anne/HDR/MemoireHDR.pdf}

\bibitem{Villadsen}
\Jorgen{} Villadsen, 
\bibtitle{Multi-dimensional Type Theory:
Rules, Categories, and Combinators for
Syntax and Semantics.}
\tinyurl{http://arxiv.org/pdf/cs/0408037.pdf} \archiveDate{15 August 2004}


\bibitem{VisettiCadiot}
Yves-Marie Visetti and Pierre Cadiot, 
\bibtitle{Instability and theory of semantic forms}.
S. Feigenbaum and D. Kurzon, eds., 
\intitle{Prepositions in their Syntactic, Semantic and Pragmatic Context}, 
Amsterdam, John Benjamins, 2002, pp. 9-39.
\tinyurl{http://formes-symboliques.org/IMG/pdf/doc-14.pdf}

\bibitem{PeiWang}
Pei Wang,
\bibtitle{What Do You Mean by \q{AI}?}
\tinyurl{https://cis.temple.edu/~pwang/Publication/AI_Definitions.pdf}


\bibitem{JacobWhitehill}
\bibtitle{Understanding ACT-R --- an Outsider's Perspective}
\tinyurl{https://arxiv.org/pdf/1306.0125.pdf}


\bibitem{WiegandMereology}
Olav K. Wiegand, 
\bibtitle{A Formalism Supplementing Cognitive Semantics Based on a New Approach to Mereology}.
Ingvar Johansson, Bertin Klein and Thomas Roth-Berghofer, eds.,  
\intitle{Contributions to the Third International Workshop on Philosophy and Informatics}, 2006. 

\bibitem{WiegandGestalts}
\underlines
\bibtitle{On referring to Gestalts}.
Mirja Hartimo, ed., \intitle{Phenomenology and Mathematics}, pp. 183-211.  Springer, 2010. 


\bibitem{Wille}
Rudolf Wille, \bibtitle{Conceptual Graphs and Formal Concept Analysis}.
\intitle{Conceptual Structures: Fulfilling Peirce's Dream
Fifth International Conference on Conceptual Structures}, 1997.


\bibitem{XiaPalmer}
Fei Xia and Martha Palmer, 
\bibtitle{Converting Dependency Structures to Phrase Structures}.
\tinyurl{http://www.aclweb.org/anthology/H01-1014}


\bibitem{YiyuYao}
Yiyu Yao, 
\bibtitle{A Comparative Study of Formal Concept
	Analysis and Rough Set Theory in Data Analysis}.
\tinyurl{http://www2.cs.uregina.ca/~yyao/PAPERS/Rough_concept.pdf}


%\bibitem{ZlatevPhenomenology}
%Jordan Zlatev, 
%\bibtitle{Phenomenology and Cognitive Linguistics}.  
%Shaun Gallagher and Dan Schmicking, eds., \intitle{Handbook on Phenomenology and Cognitive
%Science}, pp. 415-446

\bibitem{JordanZlatev}
Jordan Zlatev, 
%\underlines
\bibtitle{The Dependence of Language on Consciousness}.
\tinyurl{http://www.mrtc.mdh.se/~gdc01/work/ARTICLES/2014/4-IACAP\%202014/IACAP14-GDC/pdf/Language-Consciousness-Zlatev.pdf}

\end{thebibliography}
