\documentclass[10pt,twoside]{book}
\usepackage{fancyhdr,geometry,keyval,ifthen,makeidx,multicol,mparhack,poemscol}
\usepackage{times}
\geometry{textwidth=4in,textheight=6.75in}
\makeverselinenumbers
\global\includeaccidentalstrue
\global\includetypescriptstrue
\global\redundantemendationsfalse
\stanzaatbottom{*}
\nostanzaatbottom{\relax}
\reversemarginpar

\newcommand{\mypageref}[1]{\pageref{#1}}

\usepackage{hyperref}


\begin{document}
\maketextnotes
\makeemendations
\makeexplanatorynotes
\resetpagestyle
\normalsize
\raggedbottom
\leftheader{KADDISH FOR GAZA}
\rightheader{KADDISH FOR GAZA}
\setcounter{lineindexrepeat}{4}
\begin{volumetitlepage}
\volumetitle{KADDISH FOR GAZA}
\volumeheader{KADDISH FOR GAZA}
\volumesubtitle{by Michael Lesher}
\volumededication{There is a relationship between war and words; there is a relationship between love and war.}
\end{volumetitlepage}
\putpoemcontents
\pagenumbering{arabic}
\makepoemcontents


\poemtitle{THINKING OF GAZA FROM THOUSANDS OF MILES AWAY}
\begin{poem}
\begin{stanza}
If I were to be born again,\verseline
it might be as the white moth\verseline
whose slow, flexing heartbeat\verseline
of wings
\end{stanza}

\begin{stanza}
declares its pittance\verseline
in a vast anonymity of snow ...
\end{stanza}

\begin{stanza}
It might be as the spore\verseline
borne on mimosa tendrils\verseline
in an uncertain breeze,
\end{stanza}

\begin{stanza}
alone in a tiny pivot of air,\verseline
all the earth a mystery hovering below.
\end{stanza}

\begin{stanza}
For I am my place, and\verseline
I have nowhere to go,
\end{stanza}

\begin{stanza}
and all hearts are my heart,\verseline
and none knows me,\verseline
every breath shakes my world\verseline
though not a syllable is mine ---
\end{stanza}

\begin{stanza}
not a glance my glance, yet in every one\verseline
I disappear behind its silence ...
\end{stanza}

\begin{stanza}
And where a petal drops onto the fixed eyes\verseline
of the child whose bracts are\verseline
already in earth,\verseline
whose face\verseline
is cold with death, her eyes
\end{stanza}

\begin{stanza}
blue and blank\verseline
as the flower that was ---
\end{stanza}

\begin{stanza}
where,\verseline
at dawn, a boy rises from another tear\verseline
to test the blue air\verseline
left behind by the bomb,
\end{stanza}

\begin{stanza}
and cannot find any path\verseline
to a door, or womb, or nightmare\verseline
clearer than smoke\verseline
or louder than a shroud,
\end{stanza}

\begin{stanza}
and not even the first word\verseline
of mourning can be said ---
\end{stanza}

\begin{stanza}
I also rise;
\end{stanza}

\begin{stanza}
That is me.
\end{stanza}

\begin{stanza}
Because I am not dead.\verseline
Because I am not there.
\end{stanza}

\begin{stanza}
Because I cannot breathe\verseline
the air motionless forever\verseline
in the child's breast,
\end{stanza}

\begin{stanza}
and cannot touch the sky\margrefemendation\verseline\emendation{Ronald Reagan}
that is all that's left\verseline
to the last boy's famished eyes.
\end{stanza}

\begin{stanza}
Because I wander that sky unseen,\verseline
never to touch their earth.
\end{stanza}

\begin{stanza}
And because,\verseline
whatever I touch,\verseline
it is their faces I will feel,\verseline
their silence my breath will trace.
\end{stanza}

\end{poem}


\newpage
\poemtitle{BURLESQUE AT THE BARRIER}
\begin{poem}
\begin{stanza}
What do you write on a wall?\verseline
What can you smear that will fall\verseline
upwards of your fist, and sting\verseline
sharper than a tear?
\end{stanza}
\end{poem}


\putemendations

\putexplanatory

\puttextnotes




   \end{document}
