\spsubsectiontwoline{The co-framing system and the
doxa system}
\p{Illucutionary acts expressly signify our desire for 
something to change in our environment (with the 
help of our addressees), but similar implications 
of pragmatic desire are evident even when sentences 
are more directly assertorial, or less directly 
illocutionary.  Compare between:
\begin{sentenceList}\sentenceItem{} \label{itm:store} Remember that wine we tasted on the Niagara 
Peninsula last summer?  Can you find it in our 
local liquor store?
\sentenceItem{} \label{itm:varietal} Remember that wine we tasted on the Niagara 
Peninsula last summer?  What varietal was that again?
\end{sentenceList}
The first sentence in each pair attempts to 
establish a common frame of reference between 
addresser and addressee \mdash{} it does not, in and 
of itself, request any practical (extramental) action.  
The second sentence in (\ref{itm:store}) \i{can} be read as 
requesting that the addressee buy a bottle, though an alternate
interpretation is to learn for \i{future reference} 
whether someone \i{could} buy that bottle.  The 
second sentence in (\ref{itm:varietal}) carries no directive 
implicature at all, at least with any directness; 
it asks for more information. 
}
\p{Despite these variations, it seems reasonable to say that 
language is always performed in an overarching setting 
where concrete (extralinguistic) activity 
will \i{evetually} take place.  If in (\ref{itm:varietal}) I intend 
to recommend that grape variety to a friend, I may not be  
making a direct request of him, but I \i{am}  
proposing an eventual action that he 
might take.  If in (\ref{itm:store}) I am not issuing a directive, I 
am however establishing (and reserving the future possibility) 
that such a directive would be reasonable.  As a result, some 
extralinguistic state change seems to be lurking 
behind the linguistic content: I want my friend 
to go from having never tasted that varietal 
to having tasted it. 
Or I want to go from not having a bottle of that 
wine to having one.  Or, if I do not 
want these things at the moment, I want 
to confirm intellectually that these wishes are 
plausible.  We seem to use language to 
set up the interpersonal understandings needed 
to \i{eventually} engage in (usually collective) 
practical activity, which means effectuating some 
(extralinguistic) change.  
}
\p{That is, most expressions are not direct 
requests or suggestions of the \q{close the window} 
or \q{let's get some wine} variety, but they are 
stiches in the thread of coorinated human actions.  
Often however we use language to \i{prepare}, 
\i{negotiate}, and \i{decide upon} joint actions.  
We may have a
\i{holistic} sense that meanings orbit around 
extralinguistic and extramental state-change, 
but at the level of particular sentences most 
changes that occur, or are proposed, tend to be 
changes in our conceptualization of situations.  
Accordingly, we can pursue a semantic theory 
based on \i{change of state} if we accept that 
such changes run the gambit from changes 
\i{internal} to language \mdash{} to conversants' 
appraisal of dialogic context \mdash{} to 
changes effectuated by human activity inspired by language. 
Dialogs 
themselves change: the first sentences in 
(\ref{itm:store}) and (\ref{itm:varietal}) 
modify the discursive frame so that, for example, 
a particular wine becoms available as the anaphoric 
target for \i{that} and \i{that wine} \mdash{} and also, 
metonymically, \i{that varietal}, \i{that grape}, 
\i{that winery}.  Conceptual frames can change: 
if we are discussing a visit to Ontario and 
I mention one specific winery, one effect is to 
(insofar as the conversation follows my lead) 
refigure our joint framing to something 
narrower and more granular that the prior frame (but 
still contained in it; I am not changing the subject 
entirely).  We can pull a frame out as well as in
\mdash{} e.g., switch from talking about one winery visit to 
the whole trip, or one Leafs game to the entire season.
Moreover, our beliefs can change/evolve: if you tell me 
the wine was Cabernet Franc, I have that piece of 
info in my arsenal that I did not have before.    
}
\p{So I will argue and assume in this paper that
linguistic manings are grounded in state-changes, 
with the stipulation that the \q{register} where the changes occur can 
vary over several cognitive and extramental options: 
actual change in our environment (the window closed, 
milk in the coffee, the bottle opened); changes to the 
dialog structure (for anaphoric references, pronoun 
resolution, metalinguistic cues like \i{can you say that 
again}, etc.); changes to conceptual framings
(zoom in, zoom out, add detail); changes to beliefs.
Each of these kinds of changes deserve their own analysis, 
but we can imagine the totality of such analyses 
forming an umbrella theory of meanings. 
}
\p{During the course of a conversation \mdash{} and indeed 
of any structured cognitive activity \mdash{} we 
maintain conceptual frames representing relevant 
information; what other people know or believe;
what are our goals and plans (individually and 
collectively); and so forth.  We update
these frames periodically, and use language to 
compel others to modify their frames in
ways that we can (to some approximation) anticipate 
and encode in linguistic structure.
}
\p{In the simplest case, we can effectuate changes in 
others' frames by making assertions they are likely 
to believe to be true (assuming they deem us 
reliable).  In general, it is impossible 
to extricate the explicit content of the relevant 
speech-acts from the relevant cognitive, linguistic, and 
real-world situational contexts:
\begin{sentenceList}\sentenceItem{} That wine was a Cabernet Franc.
\sentenceItem{} Those dogs are my neighbor's.  
They are very sweet.
\end{sentenceList}
Although there is a determinate propositional content being 
asserted and although there is no propositional attitude 
other than bald assertion to complicate the pragmatics, 
still the actual words depend on addressees drawing from 
the dialogic conext in accord with how I expect them to 
(as manifest in open-ended expressions like \i{that wine}, 
\i{those dogs}, \i{they}).  Moreover, the  
open-ended components can refer outward in different 
\q{registers}: in \i{that wine} I may be 
referencing a concept previously established in 
the conversation, while \i{those dogs} may refer to 
pets we saw or heard but had not previously 
talked about.  Of course, the scenarios 
could be reversed: I could introduce \i{that wine} 
into the conversation by gesturing to a bottle 
you had not noticed before, and refer via 
\i{those dogs} to animals you have never seen or heard but 
had talked about, or heard talk about, in the recent past.
These dialog steps need to be resolved via a mixture of 
linguistic and extra-linguistic cues: 
surface-level language is not always clear as to whether 
referring expressions are to work \q{deictically} 
(drawing content from the ambient context, 
signified by gestures, rather than from any 
linguistic meaning proper), \q{discursively} 
(referring within chains of dialog, e.g. 
anaphora), or \q{descriptively} (using 
purely semantic means to establish a designation, 
like \q{my next-door neighbor's dogs} or 
\q{Inniskillin Cabernet Franc Icewine 2015}).   
}
\p{Let's agree to call the set of entities sufficiently 
relevant to a discourse or conversation context the \i{ledger}.
By \q{sufficiently relevant} I mean whatever is already 
established in a discourse so it can be referenced with something 
less that full definite description (and without the aid 
of extralinguistic gestures).  I assume that gestures
and/or descriptions are communicative acts which \q{add} 
to the ledger.  The purely linguistic case \mdash{} let's say, 
\i{descriptive additions} \mdash{} can themselves be distinguished
by their level of grounding in the current context.  
A description can be \q{definitive} in a specific situation without 
being a \i{definite description} in Russell's sense
(see \q{that wine we tasted last summer}). 
}
\p{So, descriptive additions to the ledger are one kind of 
semantic side-effect: we can change the ledger via 
language acts.  I will similarly dub another facet of 
cognitive-linguistic frames as a \i{lens}: the idea
that in conversation we can \q{zoom} attention 
in and out and move it around in time.  \q{That wine we 
tasted last summer in Ontario} both modifies 
the \i{ledger} (adding a new referent for convenient
designation) and might alter the \i{lens}:
potentially compelling subsequent 
conversation to focus on that time and/or place.  
Finally, I will identify a class of frame-modifications 
which do directly involve propositional content: 
the capacity for language to promote shared beliefs 
between people whose cognitive frames are in 
the proper resonance, by adding dtails to conceptual 
pictures already established: \i{those dogs are Staffordshires},
\i{that wine is Cabernet Franc}, \i{we have almond milk}, etc.
}
\p{For sake of discussion, I will call this latter part of the \q{active} 
cognitive frame, for some discussion 
\mdash{} the part concerning shared beliefes or asserted facts \mdash{} 
the \i{doxa inventory}.
This \q{database}-like repository stands alongside the 
\q{ledger} and \q{lens} to track propositional content 
asserted, collectively established, or already 
considered as background knowledge, \visavis{} some 
discourse.  Manipulations of the lens and ledger allow 
spakers to designate (using referential cues 
that could be ambiguous out-of-context) 
propositional contents which they 
wish to add to the \q{doxa inventory}.  I'll also say 
that modifying this inventory \i{can} be done through 
language, but participants in a discourse are entitled 
to assume that everyone formulates certain beliefs 
which are observationally obvious, and can therefore 
be linguistically presupposed rather than reported 
(the likes of that a traffic light 
is red, or a train has pulled into a station, or 
that it's raining).  
}
\p{So, I will assume that the machinery of frames is cognitive, 
not just linguistic.  We have analogous faculties for 
\q{refocusing} attention and adding conceptual details
via interaction with our environment, both alone and with others, 
and both via language and via other means.  Some 
aspects of \i{linguistic} cognitive framing \mdash{} like 
the \q{ledger} of referents previously established in a 
conversation \mdash{} may be of a purely linguistic character, 
but these are the exception rather than the rule.  
In the typical case we have a latent ability to 
direct attention and form beliefs by 
direct observation \i{or} by accepting others' reports as
proxies for direct observation.
}
\p{When we are told that two dogs are male, for instance, 
we may not perceptually encounter the dogs but we understand what 
sorts of preceptual disclosures could 
serve as motivation for someone believing that idea.  We therefore 
assume that such belief was initially warranted by 
observation and subsequently got passed through a chain 
of language-acts whose warrants are rooted in the perceived 
credibility of the speaker.  Internal to this 
process is our prior knowledge of the parameters for judging 
statements like \i{this dog is male} observationally.
}
\p{True, sometimes such observational warrants are less on 
display.  If I had never heard of Staffies (Staffordshire
pit bulls), I would be fuzzier about observational warrants
and could end up in conversations like:
\begin{sentenceList}\sentenceItem{} Those dogs are Staffordshires.
\sentenceItem{} What's a Staffordshire?
\sentenceItem{} It's a breed of dog.
\end{sentenceList}
 
\noindent{}Here I still don't really have a picture of what it is 
like to tell observationally that a dog is a Staffordshire.  
There may not be any visual cues \mdash{} at least none I 
know of \mdash{} which announce to the world that some dog's
a Staffy (compared to those announcing that it is male,
say).  But insofar as I am acquainted 
with the concept \i{dog breed}, I also understand 
the general pattern of these observations.  For instance 
I may know breeds like poodles or huskies and be able to 
identify \i{these} by distinctive visual cues.  I also 
understand that dogs' parentage is often documented, allowing 
informed parties to know their breeds via those of their 
forebearers.  That is, I am familiar with 
how beliefs about breeds are formed based on 
observation rather than just accepting others' 
reports, so I know the extralinguistic epistemology 
anchoring chains of linguistic reports in this area 
to originating observations \mdash{} even if 
I cannot in this case initiate such a chain myself.
}
\p{My overall point is that language enables us to formulate beliefs 
based on the beliefs of others, but this is possible because 
we also realize what it is like to formulate \i{our own} 
beliefs, and envision that sort of practice at the 
origin of reports that later get circulated via language.  
If we can't sufficiently picture the originating 
observations, we don't feel like we are grasping 
the linguistic simulacrum of those reports with enough 
substance.  If I never learn what Stafforshire is, 
an assertion that some dogs are Staffordshires 
has no real meaning for me \mdash{} even if I trust the asserter 
and do indeed thereby believe that the dogs are Staffordshires.  
Notice that merely knowing Staffordshire is a breed of 
dog does not expand my conceptual repertoire very much
\mdash{} it does not tell me how to recognize a Staffordshire 
or what I can do with the knowledge 
that a dog is one (it cannot, for instance, help 
me anticipate his behavior).  Nevertheless even (only) knowing 
that Staffordshire is a breed of 
dog seems to fundamentally change the status 
of sentences like \i{those dogs are Staffies}
for me: I do not \i{have} the conceptual machinery 
to exploit that knowledge, but I understand
what \i{sort} of machinery is involved.  
}
\p{In short, the \i{linguistic} meaning of concepts is tightly bound 
to how concepts factor in perceptual observations anterior 
to linguistic articulation.  As a result, 
during any episode wherein conversants use language to 
compel others' beliefs, an intrinsic dimension of the 
unfolding conversation is that people will form 
their own (extralinguistic) beliefs \mdash{} and can also 
imagine themselves in the role of originating the 
reports they hear via language, whether or not they 
can actually test out the reports by their own 
observations.
}
\p{This extralinguistic epistemic 
capacity is clearly exploited by the form of language itself.  
If a tasting organizer hands me a glass and says 
\q{This is Syrah}, she clearly expects me to infer that I 
should take the glass from her and taste the wine (and 
know that the glass contains wine, etc.).  These conventions
may be \i{mediated} by language \mdash{} we are more likely to 
understand \q{unspoken} norms by asking questions, until we gain 
enough literacy in the relevant practical domain to 
understand unspoken cues and assumptions.  But many situational 
assumptions are extralinguistic because 
they are (by convention) not explicitly stated, 
even if they accompany content that \i{is} explicitly stated.  
\i{This is Syrah} acccompanied by the gesture of handing 
me a glass is an indirect invitation for me to drink it 
(compare to \i{Please hold this for a second?} or 
\i{Please hand this to the man behind you?}). 
}
\p{I bring to every linguistic situation a 
capacity to make extralinguistc observations, and 
to understand every utterance in the context of hypothetical 
extralinguistc observations from which is originates.  
My conversation peers can use language to trigger 
these extralinguistic observations.  Sometimes the 
\q{gap} \mdash{} the conceptual slot which 
extralinguistic reasoning is expected to fill 
\mdash{} is direcly expressed, as in \i{See the 
dog over there?}.  But elsewhere the 
\q{extralinguistic implicature} is more indirect, 
as in \i{This is Syrah} and my expected belief that 
I should take and taste from the glass.  But in any 
case the phenomenon of triggering these 
extralinguistic observation is 
one form of linguistic \q{side effect}, initiating a 
change in my overall conceptualization of a situation by 
compeling me to augment beliefs with new observations. 
}
\p{All told, then, the language which is presented to me has the effect 
of initiating changes in what I believe 
\mdash{} partly via signifying propositional
content that I could take on faith, but partly 
also via directing my attention and my interpretive 
dispositions to guide me towards extralinguistic 
observations.  Here I will argue that side-effects like 
these are not side-effects \i{of} linguistic meaning, 
but are in some sense \i{constitutive} of 
meaning.  
}
\subsection{Signifying Propositional Content}
\p{Let us agree that \mdash{} beneath surface-level
co-framing complexity \mdash{} many language acts have a
transparent content as \q{doxa} that gets conveyed between
people with sufficiently resonant
cognitive frames.  So \i{in the overall course of communication} 
we have propositional contents that converge among discourse 
partners, suspended between the various cognitive and pragmatic 
units which contextualize a given, unfolding dialog.  
There is in short a \i{holistic} mapping between units of 
discourse and \q{units} of propositionality, or \q{doxa}.  
This general observation leaves unstated, however, 
\i{how} language elements map to corresponding doxic 
particulars.  I will argue that focusing on the \i{logical 
structures} of propositions can lead us astray if we 
seek to find concordant formations on the language side.
}
\p{Consider our attempts to close grandma's kitchen window.  
My analysis related to conceptual \q{transforms}
assumed that we can find, substituting for \i{literal}
propositional content, some \i{other}
(representation of a) proposition that fulfills a
speaker's unstated \q{real} meaning.  Sometimes
this makes sense: the proposition \q{that the
\i{bedroom} window is closed} can neatly,
if the facts warrant, play the role of the
proposition that \i{the kitchen window is closed}.
But we can run the example differently: there
may be \i{no} window open, but instead a draft
caused by non-airtight windows (grandma might ask
us to put towels by the cracks).  Maybe there is
no draft at all (if grandma is cold, we can
fetch her a sweater).  Instead of a single
transform, we need a a system of potential transforms
that can adapt to the facts as we discover them.
Pragmatically, the underlying problem is
that \i{grandma is cold}.  We can address this
\mdash{} if we want to faithfully respond to her request,
playing the role of cooperative conversation
partners (and grandkids) \mdash{} via a matrix
of logical possibilities:
\begin{sentenceList}\sentenceItem{} If the kitchen window is closed,
we can see if other windows are open.
\sentenceItem{} If no windows are open,
we can see if there is a draft through the window-cracks.
\sentenceItem{} If there is no draft, we can
ask if she wants a sweater.
\end{sentenceList}
This is still a logical process: starting from an
acknowledged proposition (grandma is cold) we
entertain various other propositional possibilities,
trying to rationally determine what pragmas we
should enact to alter that case
(viz., to instead make true the proposition that
\i{grandma is warm}).  Here we are not just testing
possibilities against fact, but strategically
acting to modify some facts in our environment.
}
\p{But the kind of reasoning involved here is not logical
reasoning per se: abstract logic does not tell us
to check the bedroom window if the kitchen window is closed,
or to check for gaps and cracks if all windows
are closed.  This all solicits practical, domain-specific knowledge
(about windows, air, weather, and houses).
Yet we are still deploying our practical
knowledge in logical ways \mdash{} there is a logical
structure underpinning grandma's request and
our response to it.  In sum: we (the grandkids)
are equipped with some practical knowledge
about houses and a faculty to logically utilize this
knowledge to solve the stated problem, reading
beyond the \i{explicit} form of grandma's discourse.
We use a combination of logic and background
knowledge to reinterpret the discourse as needed.
By making a request, grandma is not expressing
one attitude to one proposition, so much as
\i{initiating a process}.  This is why it would
be impolite to simply do no more
if the kitchen window is closed: our conversational
responsibility is to enact a process trying to
redress grandma's discomfort, not to entertain the
truth of any one proposition.
}
\p{For all that, there is still an overarching logical
structure here that language clearly marshals.  We read
past grandma's explicit request to infer what she is
\q{really saying} \mdash{} e.g., \i{that she is cold} \mdash{} but we
still regard her speech act in terms of its (now indirect)
propositional content.  However, notice 
how our ascertaining this content only one step toward 
legitimate understanding of the original speech-act 
(even accounting for its illocutionary dimensions).  
The doxa are \i{factors} in understanding but, 
given these cases, are not straightforward \i{designata} 
of linguistic compounds.  This implies a critique of 
truth-theoretic paradigms from a semiotic and 
compositional perspective: language is not \i{composed} 
to convey doxa through semantic reference and 
grammatic form internally (without the mediation of 
extralinguistic cognition); propositional content does 
not \i{fall out} of syntax and semantics.  I will expand 
on this critique in the next section.  
}
\p{To summarize my current arguments, then, I believe most sentences 
have an accompanying propositional content, and that during conversations 
we interpret this content as a factor in sentence meanings, becoming aware of 
what our partners believe, desire, or inquire to be the case.  We retain this 
awareness in a cumulative model of conversational context \mdash{} a \q{doxa inventory} 
\mdash{} alongside other referential and deictic axes establishing each dialogic 
setting.
}
\p{For the remainder of this paper I grant that this doxic layer is central to 
linguistic performance in general \mdash{} but given this very centrality I 
will argue that the logical substratum of language cannot be \i{separated} 
from the totality of syntactic, semantic, and pragmatic processing such 
that models based on formal logic could be curated in isolation from 
the overarching interconnectedness of language as a cognitive system.   
}
