\vspace{-2em}
\begin{thebibliography}{99}
{\fontfamily{lmtt}\selectfont\scriptsize
\vspace{-2em}
\bibitem{MartinAbadi}{%
Martin Abadi and Luca Cardelli,
\cq{A Semantics of Object Types}.  % 
Proceedings of the IEEE Symposium on Logic in Computer Science, 
Paris, 1994.
\biburl{http://lucacardelli.name/Papers/PrimObjSemLICS.A4.pdf}}


\bibitem{KennethAnderson}{%
Kenneth R. Anderson,
\cq{Freeing the Essence of a Computation}.  % 
\biburl{http://repository.readscheme.org/ftp/papers/kranderson_essence.pdf}}


\bibitem{AnglesGuttierez}{%
Renzo Angles and Claudio Guttierez,
\cq{Querying RDF Data from a Graph Database Perspective}.  % 
2005.
\biburl{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.70.225&rep=rep1&type=pdf}}


\bibitem{RonaldAshri}{%
Ronald Ashri, \i{et. al.},
\cq{Torwards a Semantic Web Security Infrastructure}.  % 
American Association for Artificial Intelligence, 2004.
\biburl{https://eprints.soton.ac.uk/259040/1/semantic_web_security.pdf}}
 
\bibitem{TimBernersLee}{%
Tim Berners-Lee,
\cq{N3Logic: A Logical Framework For the World
Wide Web}.  % 
2007.
\biburl{https://arxiv.org/pdf/0711.1533.pdf}}

\bibitem{EdwinBradyImpl}{%
Edwin Brady,
\cq{Idris, a General Purpose Dependently Typed
Programming Language: Design and
Implementation}.  % 
2013.
\biburl{https://pdfs.semanticscholar.org/1407/220ca09070233dca256433430d29e5321dc2.pdf}}

\bibitem{RBrown}{%
R. Brown, \i{et. al.},
\cq{Graphs of Morphisms of Graphs}.  % 
\biburl{https://www.emis.de/journals/EJC/Volume_15/PDF/v15i1a1.pdf}}

\bibitem{Christiansen}{%
David Raymond Christiansen, 
\cq{Practical Reflection and Metaprogramming for Dependent Types}.  % 
Dissertation, IT University of Copenhagen, 2015.
\biburl{http://davidchristiansen.dk/david-christiansen-phd.pdf}}

\bibitem{MadalinaCroitoru}{%
Madalina Croitoru and Ernesto Compatangelo,
\cq{Ontology Constraint Satisfaction Problems using Conceptual Graphs}.  % 
\biburl{https://pdfs.semanticscholar.org/d05e/eb82298201d6fae0129c6d53fe16db6d4803.pdf}}

\bibitem{ErnestoDamiani}{%
Ernesto Damiani, \i{et. al.},
\cq{Modeling Semistructured Data by Using Graph-based Constraints}.  % 
\biburl{http://home.deib.polimi.it/schreibe/TeSI/Materials/Tanca/PDFTanca/csse.pdf}}

%\bibitem{PDlotko}{%
%P. Dlotko, \i{et. al.},
%\cq{Distributed Computation of Coverage in Sensor 
%Networks by Homological Methods}.  % 
%\biburl{https://www.math.upenn.edu/~ghrist/preprints/distributedhomology.pdf}}


\bibitem{GabrieldosReis}{%
Gabriel dos Reis and Jaako J\"arvi,
\cq{What is Generic Programming?}.  % 
\biburl{https://pdfs.semanticscholar.org/e730/3991015a041e50c7bdabbe4cb4678531e35b.pdf}}


\bibitem{RichardEisenberg}{%
Richard A. Eisenberg,
\cq{Dependent Types in Haskell: Theory and Practice}.  % 
Dissertation, University of Pennsylvania, 2017.
\biburl{http://www.cis.upenn.edu/~sweirich/papers/eisenberg-thesis.pdf}}

\bibitem{ivory}{%
Trevor Elliott, \i{et. al.}
\cq{Guilt Free Ivory}.
Haskell Symposium 2015  % 
\biburl{https://github.com/GaloisInc/ivory/blob/master/ivory-paper/ivory.pdf?raw=true}}



\bibitem{MichaelEngel}{%
Michael Engel, \i{et. al.},
\cq{Unreliable yet Useful -- Reliability Annotations for Data in
Cyber-Physical Systems}.
\biburl{https://pdfs.semanticscholar.org/d6ca/ecb4cd59e79090f3ebbf24b0e78b3d66820c.pdf}}

\bibitem{MartinErwig}{%
Martin Erwig,
\cq{Specifying Type Systems with Multi-Level Order-Sorted Algebra}.
Information and Computation 207 (2009), pp. 411-437.
Third International Conference on Algebraic Methodology and Software, 
1993, pp. 177-184.
\biburl{https://web.engr.oregonstate.edu/~erwig/papers/MultiLevelAlgebra_AMAST93.pdf}}


\bibitem{MartinEscardo}{%
Mart\'{\OldI}n Escard\'o and Weng Kin Ho,
\cq{Operational domain theory and topology of sequential programming languages}.
Information and Computation 207 (2009), pp. 411-437.
\biburl{https://ac.els-cdn.com/S0890540108001570/1-s2.0-S0890540108001570-main.pdf?_tid=94a23ca4-2048-44f5-8b28-50e885faaa40&acdnat=1533048081_6911dea49597f7c7e184e5e30ae3e773f}}


\bibitem{Fabrikant}{%
Sara Irina Fabrikant,
\cq{Visualizing Region and Scale in InformationSpaces}.
Proceedings, The 20th International Cartographic Conference, ICC 2001, 
Beijing, China, Aug. 6-10, 2001, pp. 2522-2529.
\biburl{https://pdfs.semanticscholar.org/526a/09e4767ff634c4cfbc51e6f7f4ebb700096a.pdf}}


\bibitem{KathleenFisher}{%
Kathleen Fisher, \i{et. al.},
\cq{A Lambda Calculus of Objects and Method Specialization}.
Nordic Journal of Computing 1 (1994), pp. 3-37.
\biburl{https://pdfs.semanticscholar.org/5cf7/1e3120c48c23f9cecdbe5f904b884e0e1a2d.pdf}}


\bibitem{MurdochGabbay}{%
Murdoch J. Gabbay and Aleksandar Nanevski,
\cq{Denotation of Contextual Modal Type Theory (CMTT): Syntax and
Metaprogramming}.
\biburl{https://software.imdea.org/~aleks/papers/cmtt/cmtt-semantics.pdf}}


\bibitem{JeremyGibbons}{%
Jeremy Gibbons,
\cq{Datatype-Generic Programming}.
\biburl{https://www.cs.ox.ac.uk/jeremy.gibbons/publications/dgp.pdf}}


\bibitem{BenGoetzel}{%
Ben Goetzel, \i{et. al.},
\cq{Engineering General Intelligence}, Parts 1 \& 2.
Atlantis Press, 2014.
\biburl{http://wiki.opencog.org/w/Background_Publications}}


\bibitem{DineshGopalani}{%
Dinesh Gopalani, \i{et. al.},
\cq{A Type System and Type Soundness for the Calculus of 
Aspect-Oriented Programming Languages}.
Proceedings of the International MultiConference of Engineers and 
Computer Scientists (IMECS 2012) Vol 1, March 14-16, Hong Kong.
\biburl{http://www.iaeng.org/publication/IMECS2012/IMECS2012_pp263-268.pdf}}


\bibitem{MasahitoHasegawa}{%
Masahito Hasegawa,
\cq{Decomposing Typed Lambda Calculus into a Couple of 
Categorical Programming Languages}.
Proceedings of the 6th International Conference on 
Category Theory and Computer Science, 1995.
\biburl{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.53.715&rep=rep1&type=pdf}}



\bibitem{IdrisEffects}{%
Idris Development Wiki.
\cq{The Effects Tutorial}.
 % 
\biburl{http://docs.idris-lang.org/en/latest/effects/index.html}}

\bibitem{Jubien}{%
Michael Jubien,
\cq{Straight Talk about Sets}.
\i{Philosophical Topics} 17 (2), pp. 91-107 (1989)}
 % 
%\biburl{http://docs.idris-lang.org/en/latest/effects/index.html}}


\bibitem{LalanaKagal}{%
Lalana Kagal, \i{et. al.},
\cq{A Policy-Based Approached to Security for the Semantic Web}.
\biburl{https://ebiquity.umbc.edu/_file_directory_/papers/60.pdf}}


\bibitem{SunilKathari}{%
Sunil Kathari and Martin Sulzmann,
\cq{C++ templates/traits versus Haskell type classes}.
\biburl{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.78.2151&rep=rep1&type=pdf}}


\bibitem{ImanKeivanloo}{%
Iman Keivanloo, \i{et. al.},
\cq{Semantic Web-based Source Code Search}.
 % 
\biburl{http://wtlab.um.ac.ir/images/e-library/swese/Semantic\%20Web-based\%20Source\%20Code\%20Search.pdf}}



\bibitem{WernerKlieber}{%
Werner Klieber, \i{et. al.},
\cq{Using Ontologies For Software Documentation}.
 % 
\biburl{http://www.know-center.tugraz.at/download_extern/papers/MJCAI2009\%20software\%20ontology.pdf}}


\bibitem{JohnathanLee}{%
Johnathan Lee, \i{et. al.},
\cq{Task-Based Conceptual Graphs as a Basis for Automating Software Development}.  % 
\biburl{https://www.csie.ntu.edu.tw/~jlee/publication/tbcg99.pdf}}


\bibitem{HaishanLiu}{%
Haishan Liu, \i{et. al.},
\cq{Mining Biomedical Data using RDF Hypergraphs}.  % 
2013 12th International Conference on Machine Learning and Applications
\biburl{http://ix.cs.uoregon.edu/~dou/research/papers/icmla13_hypergraph.pdf}}



\bibitem{GiuseppeLongo}{%
Giuseppe Longo and Eugenio Moggi,
\cq{A Category-Theoretic Characterization of Functional Completeness}.  % 
Theoretical Computer Science, 70 (2), 1990, pp.193-211.
\biburl{https://pdfs.semanticscholar.org/58a7/62acf822fd3520eeb1576b7758351d05c34d.pdf}}

\bibitem{KatharinaMehner}{%
Katharina Mehner, \i{et. al.},
\cq{Analysis of Aspect-Oriented Model Weaving}.  % 
\biburl{http://www.mathematik.uni-marburg.de/~swt/Publikationen_Taentzer/MMT09.pdf}}

\bibitem{MarkMinas}{%
Mark Minas and Hans J Schneider,
\cq{Graph Transformation by Computational Category Theory}.  % 
\biburl{https://www2.informatik.uni-erlangen.de/staff/schneider/gtbook/fmn-final.pdf}}


\bibitem{GiladMishne}{%
GiladMishne and Maarten de Rijke,
\cq{Source Code Retrieval using Conceptual Similarity}.  % 
\biburl{https://staff.fnwi.uva.nl/m.derijke/Publications/Files/riao2004.pdf}}

\bibitem{SantanuPaul}{%
Santanu Paul and Atul Prakash,
\cq{Supporting Queries on Source Code: A Formal Framework}.  % 
\biburl{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.52.9136&rep=rep1&type=pdf}}

\bibitem{HeikoPaulheim}{%
Heiko Paulheim and Christian Bizer,
\cq{Type Inference in Noisy RDF Data}.  % 
\biburl{http://www.heikopaulheim.com/docs/iswc2013.pdf}}



\bibitem{AlexandraPoulovassilis}{%
Alexandra Poulovassilis and Mark Levene,
\cq{A Nested-Graph Model for the Representation and 
Manipulation of Complex Objects}.
Data and. Knowledge Engineering, 6, 3 (1991), pp. 205-224
\biburl{http://www.dcs.bbk.ac.uk/~mark/download/tois.pdf}}


\bibitem{LavanyaRamapantulu}{%
Lavanya Ramapantulu, \i{et. al.},
\cq{A Conceptual Framework to Federate Testbeds for Cybersecurity}.
Proceedings of the 2017 Winter Simulation Conference.
\biburl{http://simulation.su/uploads/files/default/2017-ramapantulu-teo-chang.pdf}}

\bibitem{WilliamRapaport}{%
William J. Rapaport,
\cq{Semiotic Systems, Computers, and the Mind:  How cognition could be computing}.
International Journal of Signs and Semiotic Systems, 2(1), 32-71, January-June 2012.
\biburl{https://cse.buffalo.edu/~rapaport/Papers/Semiotic_Systems,_Computers,_and_the_Mind.pdf}}


\bibitem{YuriySolodkyy}{%
Yuriy Solodkyy, \i{et. al.},
\cq{Open and Efficient Type Switch for C++}.
\biburl{http://www.stroustrup.com/OOPSLA-typeswitch-draft.pdf}}

\bibitem{JohnStell}{%
John G. Stell,
\cq{Granulation for Graphs}.
International Journal of Signs and Semiotic Systems, 2(1), 32-71, January-June 2012.
\biburl{https://pdfs.semanticscholar.org/9e0f/a93a899e36dc3df62feabc004a0ecef4365d.pdf}}



\bibitem{TakeshiTakahashi}{%
Takeshi Takahashi, \i{et. al.},
\cq{Ontological Approach toward Cybersecurity in Cloud Computing}.
3rd International Conference on Security of Information and Networks (SIN 2010), 
Sept. 7-11, 2010, Taganrog, Rostov Oblast, Russia.
\biburl{https://arxiv.org/pdf/1405.6169.pdf}}

\bibitem{MozhganTavakolifard}{%
Mozhgan Tavakolifard,
\cq{On Some Challenges for Online 
Trust and Reputation Systems}.  % 
Dissertation, Norwegian University of Science and Technology, 2012.
\biburl{https://pdfs.semanticscholar.org/fc60/d309984eddd4f4229aa56de2c47f23f7b65e.pdf}}

\bibitem{BradyMatus}{%
Mat\'u\c{s} Teji\v{s}\v{c}\'ak and Edwin Brady,
\cq{Practical Erasure in Dependently Typed Languages}.  % 
\biburl{https://eb.host.cs.st-andrews.ac.uk/drafts/dtp-erasure-draft.pdf}}

\bibitem{TillyEtAl}{%
Scott R. Tilley, \i{et. al.},
\cq{Towards a Framework for Program Understanding}.  % 
\biburl{https://pdfs.semanticscholar.org/71d0/4492be3c2abf9e1a88b9b263193a5c51eff1.pdf}}


\bibitem{TuckerZucker}{%
J. V. Tucker  and  J. I. Zucker, 
\cq{Computation by `While' Programs on Topological Partial Algebras }.  % 
Theoretical Computer Science 219 (1999), pp. 379-420. 
\biburl{https://core.ac.uk/download/pdf/82201923.pdf}}


\bibitem{TurnerEden}{%
Raymond Turner and Amnon H. Eden, 
\cq{Towards a Programming Language Ontology}.  % 
\biburl{https://www.researchgate.net/publication/242381616_Towards_a_Programming_Language_Ontology}}


\bibitem{EricWalkingshaw}{%
Eric Walkingshaw, 
\cq{The Choice Calculus: A Formal Language of Variation}.  % 
Dissertation, Oregon State University, 2013
\biburl{http://web.engr.oregonstate.edu/~walkiner/papers/thesis-choice-calculus.pdf}}

\bibitem{YurickWilks}{%
Yurick Wilks, 
\cq{The Semantic Web as the Apotheosis of Annotation, 
but What Are Its Semantics?}.  % 
\biburl{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.131.4958&rep=rep1&type=pdf}}

\bibitem{ReneWitte}{%
Rene Witte,  \i{et. al.}, 
\cq{Ontological Text Mining of Software Documents}.  % 
\biburl{https://pdfs.semanticscholar.org/7034/95109535e510f81b9891681f99bae1e704fc.pdf}}


\bibitem{PornpitWongthongtham}{%
Pornpit Wongthongtham,  \i{et. al.}, 
\cq{Development of a Software Engineering Ontology for 
Multi-site Software Sevelopment}.  % 
\biburl{https://ifs.host.cs.st-andrews.ac.uk/Research/Publications/Papers-PDF/2005-09/TKDE-Ponpit-2009.pdf}}

\bibitem{EdwardZalta}{%
Edward N. Zalta, 
\cq{The Modal Object Calculus and its Interpretation}.  % 
\biburl{https://mally.stanford.edu/Papers/calculus.pdf}}


\bibitem{CharlesZhang}{%
Charles Zhang and Hans-Arno Jacobsen, 
\cq{Refactoring Middleware with Aspects}.
IEEE Transactions on Parallel and Distributed Systems 
Vol. 14 No. 12, November 2003.   % 
\biburl{https://pdfs.semanticscholar.org/0304/5c5cc518c7d44c3f7b117ea11dfae4932a89.pdf?_ga=2.207853466.903112516.1533046888-196394048.1525384494}}

}

\end{thebibliography}