
%\vspace{.4em}

%\onecolumngrid
\newsavebox{\dlbox}
\begin{lrbox}{\dlbox}{\sectionlinenoc{darkRed!60!cyan}{0.9\linewidth}}
\end{lrbox}

{\begin{flushright}
\raisebox{-.2em}{%\hspace*{1em}
	\usebox{\dlbox}}
	\end{flushright}}

\onecolumn
\begin{thebibliography}{99}
	
\vspace{-1.5em}

%
\bibitem{RaubalAdams}
Benjamin Adams and Martin Raubal, 
\bibtitle{A Metric Conceptual Space Algebra}.
\tinyurl{https://pdfs.semanticscholar.org/521a/cbab9658df27acd9f40bba2b9445f75d681c.pdf}

\bibitem{RaubalAdamsCSML}
Benjamin Adams and Martin Raubal, 
\bibtitle{Conceptual Space Markup Language (CSML): Towards the Cognitive Semantic Web}.
\tinyurl{http://idwebhost-202-147.ethz.ch/Publications/RefConferences/ICSC_2009_AdamsRaubal_Camera-FINAL.pdf}

\bibitem{RaubalAdamsMore}
Benjamin Adams and Martin Raubal, 
\bibtitle{The Semantic Web Needs More Cognition}.
\tinyurl{http://www.semantic-web-journal.net/sites/default/files/swj37_0.pdf}


\bibitem{NathanielChristenCyberphysical}
Nathaniel Christen, 
\bibtitle{Hypergraph Type Theory for 
Specifications-Conformant Code and 
Generalized Lambda Calculus}.
\intitle{Advances in Ubiquitous Computing: Cyber-Physical Systems, Smart Cities, and Ecological Monitoring}, Amy Neustein, ed., Elsevier, 2019. 


\bibitem{InteractingConceptualSpaces}
Bob Coecke, \i{et. al.}, 
\bibtitle{Interacting Conceptual Spaces I:
Grammatical Composition of Concepts}.
\tinyurl{https://arxiv.org/pdf/1703.08314.pdf}

\bibitem{BrendanFong}
Brendan Fong, \bibtitle{Decorated Cospans}
\tinyurl{https://arxiv.org/abs/1502.00872}

\bibitem{BrendanFongThesis}
\repeatName, \bibtitle{The Algebra of Open and
Interconnected Systems}.
Oxford University, dissertation 2016.
\tinyurl{https://arxiv.org/pdf/1609.05382.pdf}

\bibitem{PeterGardenfors}
Peter \Gardenfors{},  
\bibtitle{How to Make the Semantic Web More Semantic}.
\tinyurl{https://pdfs.semanticscholar.org/1d58/faa5cb23efb7394aeb3dfe688edd99797a91.pdf}


\bibitem{GardenforsBook}
\repeatName, 
\bibtitle{Conceptual Spaces: The Geometry of Thought}.
MIT Press, 2000.


\bibitem{GardenforsZenker}
Peter \Gardenfors{} and Frank Zenker,  
\bibtitle{Theory Change as Dimensional Change: Conceptual Spaces 
	Applied to the Dynamics of Empirical Theories}.
\intitle{Synthese 190(6)}, pp. 1039-1058, 2013.  
\tinyurl{http://lup.lub.lu.se/record/1775234}

 
\bibitem{BenGoertzel}
Ben Goertzel, 
\bibtitle{A Comparison of the Novamente AI Design with the Human Mind/Brain}.
\tinyurl{http://www.goertzel.org/new_research/NM_human_psych.pdf}

\bibitem{GoertzelPLN}
\repeatName, 
\bibtitle{Probabilistic Language Networks:
Integrating Word Grammar and Link Grammar 
in the Framework of Probabilistic Logic}.
\tinyurl{http://goertzel.org/ProwlGrammar.pdf}


\bibitem{GoertzelSS}
\repeatName, 
\bibtitle{Perception Processing for General Intelligence: Bridging 
the Symbolic/Subsymbolic Gap}.
\tinyurl{http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.374.6985&rep=rep1&type=pdf}

\bibitem{GoertzelP2}
\repeatName, 
\bibtitle{Engineering General Intelligence, Part 2}.
Paris: Atlantis Press, 2014

\bibitem{GoertzelEtAl}
Ben Goertzel, \i{et. al},
\bibtitle{A General Intelligence Oriented Architecture for 
Embodied Natural Language Processing}.
\tinyurl{https://pdfs.semanticscholar.org/c125/5ab910c12cfc948889ce5149ed05ae552870.pdf}

\bibitem{HolmqvistDiss}
Kenneth Holmqvist, 
\bibtitle{Implementing Cognitive Semantics: Image schemata, 
valence accommodation, and valence suggestion for  
AI and computational linguistics}.
PhD thesis, Dept. of Cognitive Science, Lund University, Lund, Sweden, 1993.

\bibitem{WolfgangJeltsch}
Wolfgang Jeltsch, 
\bibtitle{Categorical Semantics for Functional Reactive 
Programming with Temporal Recursion and Corecursion}.
\tinyurl{https://arxiv.org/pdf/1406.2062.pdf}

\bibitem{AleksKissinger}
Aleks Kissinger, 
\bibtitle{Finite Matrices are Complete for (dagger-)Hypergraph Categories}.
\tinyurl{https://arxiv.org/abs/1406.5942}

\bibitem{RuitingLian}
Ruiting Lian, \i{et. al.}, 
\bibtitle{Syntax-Semantic Mapping for General
Intelligence: Language Comprehension as
Hypergraph Homomorphism, Language
Generation as Constraint Satisfaction}.
\tinyurl{http://www.frontiersinai.com/turingfiles/December/lian.pdf}


\bibitem{BalintMolnar}
B\'alint Moln\'ar, 
\bibtitle{Applications of Hypergraphs In informatics: A survey and 
opportunities for research}
\tinyurl{http://ac.inf.elte.hu/Vol_042_2014/261_42.pdf}

\bibitem{JenniferPaykin}
Jennifer Paykin, et. al.,
\bibtitle{Curry-Howard for GUIs: 
Or, User Interfaces via Linear Temporal, Classical Linear Logic}.
\tinyurl{https://www.cl.cam.ac.uk/~nk480/obt.pdf}

\bibitem{PaykinKrishnaswami}
\repeatName, 
\bibtitle{The Essence of Event-Driven Programming}
\tinyurl{https://jpaykin.github.io/papers/pkz_CONCUR_2016.pdf}

%\bibtitle{Linear Temporal Type Theory for Event-based Reactive Programming}.
%\tinyurl{https://www.semanticscholar.org/paper/Linear-Temporal-Type-Theory-for-Event-based-Paykin-%Krishnaswami/4b8eccab1340c9d1035728ba5b198eab41ab66f3}


\bibitem{PetitotSyntaxe}
Jean Petitot, 
\bibtitle{Syntaxe Topologique et Grammaire Cognitive}
\tinyurl{https://www.persee.fr/doc/lgge_0458-726x_1991_num_25_103_1610}


\bibitem{MattSelway}
Matt Selway, \i{et. al},
\bibtitle{Configuring Domain Knowledge for Natural Language Understanding}
\tinyurl{http://ceur-ws.org/Vol-1128/paper9.pdf}


%\bibitem{SmithGibbons}
%Michael Anthony Smith and Jeremy Gibbons,
%\bibtitle{Unifying Theories of Objects}
%\tinyurl{http://www.cs.ox.ac.uk/jeremy.gibbons/publications/uto.pdf}


\bibitem{SpivakKent}
David Spivak and Robert Kent, 
\bibtitle{Ologs: A Categorical Framework for Knowledge
Representation}
\tinyurl{https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0024274&type=printable}

\bibitem{GregorStrle}
Gregor Strle, \bibtitle{Semantics Within: The Representation of Meaning Through Conceptual Spaces}.  
University of Novi Gorici, dissertation 2012.
\tinyurl{http://repozitorij.ung.si/Dokument.php?id=738&lang=slv}

\bibitem{TemperleySleator}
Davy Temperley and Daniel Sleator,
\bibtitle{Parsing English with a Link Grammar}
\tinyurl{https://www.link.cs.cmu.edu/link/ftp-site/link-grammar/LG-IWPT93.pdf}


\bibitem{BaltasarTrancon}
Baltasar Tranc'on y Widemann, \i{et. al.}, 
\bibtitle{Automized Generation of Typed Syntax Trees via XML}
\tinyurl{http://pizza.cs.ucl.ac.uk/xse01/ready/10.pdf}


\bibitem{OrlinVakarelov}
Orlin Vakarelov, 
\bibtitle{Pre-cognitive Semantic Information}
\tinyurl{http://www.u.arizona.edu/~okv/Papers/Pre-CogSemInfo_KTB.pdf}

\bibitem{MartinWesthead}
Martin Westhead, \bibtitle{BinX -- The Binary XML Description Language}
\tinyurl{http://buphy.bu.edu/~brower/SciDAC/doc/BinXIntro2.pdf}

\end{thebibliography}
